Fast Construction of Inter-Object Spacing Representations

### 2017-04-01 paper.tex
First major text revision addressing reviewers' comments and adding figures

\documentclass[submission]{gmp2017}
%\documentclass{gmp2017}

\title{Parallel Quadtree Construction on Collections of Objects}

% Put your submission number here.
% This number will appear instead of authors and affiliations if the
% class option "submission" is active.
\SubNumber{19}

% Put the author names here.
% Use the second argument as a reference to the list of affiliations.
% No authors and affiliations will appear if the class option "submission"
% is active.
\author{Nathan Vollmer}{1}
\author{John Edwards}{1}

% Put the affiliations of the authors here.
\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}

@@

%\teaser{
%  \subfloat[][]{
%    \label{fig:gears1}
%    \begin{tikzpicture}
%      \node[anchor=south west,inner sep=0] at (0,0) {
%        \begin{tabular}[b]{c}
%          \includegraphics[height=0.8in]{gears-far1.png} \\
%          \includegraphics[height=0.8in]{gears-close1.png}
%        \end{tabular}
%      };
%      \draw[black,thick] (1.4,3.2) rectangle (2.0,3.6);
%      \draw[black,dashed] (1.4,3.6) -- (0.22,2.15);
%      \draw[black,dashed] (2.0,3.6) -- (2.95,2.15);
%      \draw[black,thick] (0.22,0.1) rectangle (2.95,2.15);
%    \end{tikzpicture}
%  }
%  \subfloat[][]{
%    \label{fig:gears}
%    \begin{tikzpicture}
%      \node[anchor=south west,inner sep=0] at (0,0) {
%        \begin{tabular}[b]{c}
%          \includegraphics[height=0.8in]{gears-far4.png} \\
%          \includegraphics[height=0.8in]{gears-close4.png}
%        \end{tabular}
%      };
%      \draw[black,thick] (1.3,3.1) rectangle (1.9,3.5);
%      \draw[black,dashed] (1.3,3.5) -- (0.22,2.15);
%      \draw[black,dashed] (1.9,3.5) -- (2.92,2.15);
%      \draw[black,thick] (0.22,0.1) rectangle (2.92,2.15);
%    \end{tikzpicture}
%  }
%  \hspace{3mm}
%  \subfloat[][]{
%    \label{fig:knife}
%    \includegraphics[trim=4cm 0cm 4cm 2.5cm, clip=true, height=1.4in]
%                    {knife-above/slice-00000.png}
%  }
%  \subfloat[][]{
%    \label{fig:knife2}
%    \includegraphics[trim=2mm 0cm 2mm 2cm, clip=true, height=1.4in]
%                    {knife-above/slice-00110.png}
%  }
%  \caption{Two example applications of the \red{approximated} generalized Voronoi diagram (GVD) computed by our novel, adaptive algorithm. Previous GVD methods require a gridded space of $2^{24}$ (gears dataset) and $2^{36}$ (knives dataset) voxels to resolve the closely spaced objects.
%    \protect\subref{fig:gears1} Two gears with regions of very tight spacing.
%    \protect\subref{fig:gears} The GVD of the gears model.  The surface is colored red in areas of very close tolerance.
%    \protect\subref{fig:knife} Three butter knives in a wood block.  To animate removal of the knives without intersecting the block requires extreme care because of close mesh spacing.
%    \protect\subref{fig:knife2} Intersection-free motion is guaranteed by computing motion vectors based on the GVD and allowing motion only within a Voronoi cell.
%  }
%  \label{fig:teaser}
%}
%
\maketitle

\begin{abstract}
We present a parallel quadtree algorithm that resolves between geometric objects. Our quadtree has the property that no cell intersects more than one labeled object. A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori, because the number of cells required to resolve closely spaced objects exceeds available memory. Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation. Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time). We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D. Our results show an order of magnitude improvement over the current state of the art.

@@

@@

%-------------------------------------------------------------------------------
% introduction
%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, object description, and other applications. Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects. However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object. Using a quadtree, we can model the space between objects, the first step in constructing distance fields, detecting collisions, and computing the generalized Voronoi diagram. Modeling inter-object spacing is computationally straightforward when inter-object spacing is large compared to the world bounding box. Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.

Difficulties arise when spacing between objects is small relative to the size of the domain. An approach using a uniform grid would have excessive memory requirements in order to resolve between objects because the uniformly sized grid cell must be small enough to fit between objects at every location in the domain. Thus, an adaptive approach must be used for datasets of closely spaced objects. To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory but does so in serial, with expected performance liabilities.

This work extends the work done by Edwards et al \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm targeted for the GPU. Our algorithm performs an order of magnitude faster than the previous work and will be an excellent base for later distance transform and generalized Voronoi diagram computation.

@@

\begin{enumerate}
\item Construct an quadtree on object vertices using Karras' algorithm \cite{karras2012maximizing}
\item Detect quadtree cells that intersect more than one object, which we call ``conflict cells'' (contribution)
\item Subdivide conflict cells to resolve objects (contribution)
\end{enumerate}

@@

\paragraph{Serial} In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute octrees over a set of solid models. 
% JME: note that in the Bastos work, ``reconstruction'' is reconstruction
% of the surface from the quadtree at rendering time.
Two seminal works build octrees on objects in order to compute the Adaptive Distance Field (ADF) on octree vertices.  Strain~\shortcite{strain1999fast} fully resolves the quadtree everywhere on the object surface, and Frisken \etal~\shortcite{frisken2000adaptively} resolve the quadtree fully only in areas of small local feature size.  Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation. Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.  Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.

@@

Most of our algorithm is independent of dimension, and so we will use terms ``facet'' and ``octree'' (which lacks a dimension-independent term). The process of finding sample points to resolve conflict cells is limited to 2D in this paper, so for that section we will use ``line'' and ``quadtree''.

\subsection{Build initial octree}
\label{sec:build-initial-octree}
Our first step is to build an octree on the vertices. This step gives us our first approximation to our final octree. We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel. With the binary radix tree, the quadtree can be constructed with a single parallel call. The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points. That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.

@@

Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$, where the local address is the Morton code of a $2^{DIM}$ block of voxels. The address of the root cell is defined as the empty string. Figure \ref{fig:conflict-find-4} shows the address of each leaf cell in a quadtree.

@@

Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (Figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0,outer sep=0] at (0,0) {
      \subfloat[][]{
        \label{fig:conflict-find-1}
        \includegraphics[width=0.22\columnwidth]{conflict-find-1.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-2}
        \includegraphics[width=0.22\columnwidth]{conflict-find-2.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-4}
        \includegraphics[width=0.22\columnwidth]{conflict-find-4.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-5}
        \includegraphics[width=0.22\columnwidth]{conflict-find-5.pdf} }
    };
    %\draw[red,ultra thick] (4.0,1.7) rectangle (6.55,4.2);
    \draw[red,thick] (4.0,1.7) rectangle (6.55,4.2);

    % lines
    % upper-right
    \draw[red,thin] (6.55,4.2) -- (11.8,4.2);
    % lower-left
    \draw[red,thin] (4.0,1.7) -- (8.0,0.3);

    %% curves
    %% upper-left
    %%\draw[red,thin] (4.0,4.2) to[out=10,in=170] (8.0,4.2);
    %% upper-right
    %\draw[red,thin] (6.55,4.2) to[out=10,in=170] (11.8,4.2);
    %% lower-left
    %\draw[red,thin] (4.0,1.7) to[out=-40,in=175] (8.0,0.3);

    \draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);
  \end{tikzpicture} \\
      \subfloat[][]{
        \label{fig:conflict-find-6}
        \includegraphics[width=0.22\columnwidth]{conflict-find-6.pdf} } \hfill
      \subfloat[][]{
        \label{fig:data-structures-1}
        \includegraphics[width=0.22\columnwidth]{data-structures-1.pdf} } \hfill
      \subfloat[][]{
        \label{fig:data-structures-2}
        \includegraphics[width=0.22\columnwidth]{data-structures-3.pdf} } \hfill
      \subfloat[][]{
        \label{fig:data-structures-3}
        \includegraphics[width=0.22\columnwidth]{data-structures-4.pdf} }
  \caption{
    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
    \protect\subref{fig:conflict-find-2} We construct an initial quadtree on the vertices using Karras' algorithm.
    \protect\subref{fig:conflict-find-4} Zoomed-in view of the boundary cell (BCell) computation for each facet. These pairs are given in figure \protect\subref{fig:data-structures-1}.
    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:conflict-find}
\end{figure}

\begin{figure}
  \label{fig:scc-sort}
\end{figure}

@@

\subsection{Resolve conflict cells}

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:conflict-resolution-x}
    \includegraphics[width=0.24\columnwidth]{conflict-resolution-x.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-even}
    \includegraphics[width=0.24\columnwidth]{conflict-resolution-even.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-all}
    \includegraphics[width=0.24\columnwidth]{conflict-resolution-all.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-octree}
    \includegraphics[width=0.24\columnwidth]{conflict-resolution-octree.pdf} } \\
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-even}
    \includegraphics[width=0.24\columnwidth]{conflict-resolution-adjacent-even.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-all}
    \includegraphics[width=0.24\columnwidth]{conflict-resolution-adjacent-all.pdf} }
  \caption{
    \protect\subref{fig:conflict-resolution-x} A conflict cell with two lines from different objects.
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples. This ensures that an quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
  }
  \label{fig:conflict-resolution}
\end{figure}

>>>>

\documentclass[final,3p,times,twocolumn]{elsarticle}
%\documentclass[submission]{gmp2017}
%\documentclass{gmp2017}

% Put your submission number here.
% This number will appear instead of authors and affiliations if the
% class option "submission" is active.
%\SubNumber{19}

% Put the author names here.
% Use the second argument as a reference to the list of affiliations.
% No authors and affiliations will appear if the class option "submission"
% is active.
%\author{Nathan Vollmer}{1}
%\author{John Edwards}{1}

% Put the affiliations of the authors here.
%\affiliation{1}{Department of Informatics and Computer Science, Idaho State University, USA}

%% The numcompress package shorten the last page in references.
%% `nodots' option removes dots from firstnames in references.
\usepackage[nodots]{numcompress}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
\usepackage{lineno}

%% Avoids linenumbers to collide with text for 5p format:
\setlength\linenumbersep{3pt}

@@

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

%\title{Parallel Quadtree Construction on Collections of Objects}
\title{Fast Construction of Inter-Object Spacing Representations}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\author{}

\address{}

%\teaser{
%  \subfloat[][]{
%    \label{fig:gears1}
%    \begin{tikzpicture}
%      \node[anchor=south west,inner sep=0] at (0,0) {
%        \begin{tabular}[b]{c}
%          \includegraphics[height=0.8in]{gears-far1.png} \\
%          \includegraphics[height=0.8in]{gears-close1.png}
%        \end{tabular}
%      };
%      \draw[black,thick] (1.4,3.2) rectangle (2.0,3.6);
%      \draw[black,dashed] (1.4,3.6) -- (0.22,2.15);
%      \draw[black,dashed] (2.0,3.6) -- (2.95,2.15);
%      \draw[black,thick] (0.22,0.1) rectangle (2.95,2.15);
%    \end{tikzpicture}
%  }
%  \subfloat[][]{
%    \label{fig:gears}
%    \begin{tikzpicture}
%      \node[anchor=south west,inner sep=0] at (0,0) {
%        \begin{tabular}[b]{c}
%          \includegraphics[height=0.8in]{gears-far4.png} \\
%          \includegraphics[height=0.8in]{gears-close4.png}
%        \end{tabular}
%      };
%      \draw[black,thick] (1.3,3.1) rectangle (1.9,3.5);
%      \draw[black,dashed] (1.3,3.5) -- (0.22,2.15);
%      \draw[black,dashed] (1.9,3.5) -- (2.92,2.15);
%      \draw[black,thick] (0.22,0.1) rectangle (2.92,2.15);
%    \end{tikzpicture}
%  }
%  \hspace{3mm}
%  \subfloat[][]{
%    \label{fig:knife}
%    \includegraphics[trim=4cm 0cm 4cm 2.5cm, clip=true, height=1.4in]
%                    {knife-above/slice-00000.png}
%  }
%  \subfloat[][]{
%    \label{fig:knife2}
%    \includegraphics[trim=2mm 0cm 2mm 2cm, clip=true, height=1.4in]
%                    {knife-above/slice-00110.png}
%  }
%  \caption{Two example applications of the \red{approximated} generalized Voronoi diagram (GVD) computed by our novel, adaptive algorithm. Previous GVD methods require a gridded space of $2^{24}$ (gears dataset) and $2^{36}$ (knives dataset) voxels to resolve the closely spaced objects.
%    \protect\subref{fig:gears1} Two gears with regions of very tight spacing.
%    \protect\subref{fig:gears} The GVD of the gears model.  The surface is colored red in areas of very close tolerance.
%    \protect\subref{fig:knife} Three butter knives in a wood block.  To animate removal of the knives without intersecting the block requires extreme care because of close mesh spacing.
%    \protect\subref{fig:knife2} Intersection-free motion is guaranteed by computing motion vectors based on the GVD and allowing motion only within a Voronoi cell.
%  }
%  \label{fig:teaser}
%}
%
%\maketitle

\begin{abstract}
%We present a parallel quadtree algorithm that resolves between geometric objects. Our quadtree has the property that no cell intersects more than one labeled object. A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori, because the number of cells required to resolve closely spaced objects exceeds available memory. Previous parallel algorithms that are spatially adaptive, or that discretize finely in only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation. Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time). We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D. Our results show significant improvement over the current state of the art.
We present a parallel quadtree algorithm that resolves between geometric objects, modeling space between objects rather than the objects themselves. Our quadtree has the property that no cell intersects more than one labeled object. A popular technique for discretizing space is to impose a uniform grid -- an approach that is easily parallelizable but often fails because object separation isn't known a priori or because the number of cells required to resolve closely spaced objects exceeds available memory. Previous parallel algorithms that are spatially adaptive, discretizing finely only where needed, either spawn kernels hierarchically, separate points only, or make no guarantees of object separation. Our 2D algorithm is the first to construct an object-resolving discretization that is hierarchical (saving memory) yet with a fully parallel approach (saving time). We describe our algorithm, derive the time complexity, demonstrate experimental results, and discuss extension to 3D. Our results show significant improvement over the current state of the art.

@@

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
\linenumbers

@@

%-------------------------------------------------------------------------------
% introduction
%-------------------------------------------------------------------------------
\section{Introduction}
\label{sec:intro}
Constructing quadtrees on objects is an important task with applications in collision detection, distance fields, robot navigation, shape modeling, object description, and other applications. Quadtrees built on objects most often model the objects themselves, providing a space-efficient representation of arbitrarily complex objects. However, our work centers on using quadtrees to separate, or resolve, collections of closely spaced objects, i.e., to construct a discretization such that no cell intersects more than one object. Using a quadtree, we can model the space between objects. Our quadtree representation is not purposed for fast retrieval as is often the case in hierarchical subdivisions of point data. Rather, the quadtrees are used for constructing distance fields, detecting collisions and computing the GVD \cite{edwards2015approximating}.

\red{I just added this paragraph. What do you think?} Object separation, such as is treated here, is of some use in 2D (e.g. path planning), but becomes a very important problem in 3D. Hierarchically subdividing non-point data in a principled parallel way is surprisingly complex, and this paper lays the groundwork for our continuing work in 3D.

Modeling inter-object spacing is computationally straightforward when inter-object spacing is large compared to the world bounding box. Approaches typically involve a uniform grid of the space, which leads to efficient computation that often uses graphics processors.

Difficulties arise when objects are close together relative to the size of the domain. An approach using a uniform grid would have excessive memory requirements in order to resolve between objects because the uniformly sized grid cell must be small enough to fit between objects at every location in the domain. Thus, an adaptive approach must be used for datasets of closely spaced objects. To our knowledge, only one algorithm \cite{edwards2015approximating} computes an adaptive data structure that fully resolves between objects without using unreasonable amounts of memory, but it does so in serial, with expected performance liabilities. A naive approach to parallelizing quadtree computation would be to assign all available compute units according to a course grid, then run the serial algorithm on each compute unit. While simple, there is potential for serious load imbalancing if the close object spacings are not uniformly distributed.

This work extends the work done by Edwards \etal \cite{edwards2015approximating} by computing the quadtree in parallel with an algorithm that is adaptive and independent of object distribution. Our algorithm, which is targeted for the GPU, performs an order of magnitude faster than the previous work and will be an important base for later distance transform and generalized Voronoi diagram computation.

@@

\begin{enumerate}
\item Construct a quadtree on object vertices using the Karras algorithm \cite{karras2012maximizing}
\item Detect quadtree cells that intersect more than one object, which we call ``conflict cells'' (contribution)
\item Subdivide conflict cells to resolve objects (contribution)
\end{enumerate}

@@

\paragraph{Serial} In an early work, Lavender \etal \shortcite{lavender1992voronoi} define and compute octrees over a set of solid models. 
% JME: note that in the Bastos work, ``reconstruction'' is reconstruction
% of the surface from the quadtree at rendering time.
Two seminal works build octrees on objects in order to compute the Adaptive Distance Field (ADF) on octree vertices.  Strain \shortcite{strain1999fast} fully resolves the quadtree everywhere on the object surface, and Frisken \etal \shortcite{frisken2000adaptively} resolve the quadtree fully only in areas of small local feature size.  Both approaches are designed to retain features of a single object rather than resolving between multiple objects, as is required for GVD computation. Boada \etal \shortcite{boada2002voronoi,boada2008approximations} use an adaptive approach to GVD computation, but their algorithm is restricted to GVDs with connected regions and is inefficient for polyhedral objects with many facets.  Two other works are adaptive \cite{teichmann1997polygonal,vleugels1998approximating} but are computationally expensive and are restricted to convex sites.

@@

%Most of our algorithm is independent of dimension, and so we will use terms ``facet'' and ``octree'' (which lacks a dimension-independent term). The process of finding sample points to resolve conflict cells is limited to 2D in this paper, so for that section we will use ``line'' and ``quadtree''.
Each step of our algorithm, with the exception of resolving conflict cells, is independent of dimension and can be used for 3D octree applications. But since point sampling for conflict cell resolution is 2D we will use the term quadtree throught the algorithm description for consistency. Our algorithm assumes the objects are faceted where the facets are simplices.

\subsection{Build initial octree}
\label{sec:build-initial-octree}
Our first step is to build an octree on the given set of vertices. We use the Karras algorithm \cite{karras2012maximizing} which starts by sorting the Morton codes of the given vertices. Our implementation uses an efficient parallel radix sorter described by Ha \etal \shortcite{ha2009fast}. Once the vertices are sorted, a binary radix tree, and then an initial octree can be constructed in parallel. The strength of this approach lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points. That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.

%Our first step is to build an octree on the vertices. This step gives us our first approximation to our final octree. We use Karras' algorithm \cite{karras2012maximizing} which sorts the Morton codes of the vertices in parallel, then constructs the binary radix tree in parallel. With the binary radix tree, the quadtree can be constructed with a single parallel call. The strength of this algorithm lies in the fact that overall performance scales linearly with the number of cores, regardless of the distribution of points. That is, even if a large number of vertices are clustered in a small area, requiring deep quadtree subdivision, only a constant number of parallel calls need be made.

\subsection{Pruning the octree}
Assume we have a vertex labeling, such that each vertex is labeled with the object it belongs to. The initial octree resolves between vertices, regardless of whether they are labeled differently from each other or not, so each leaf node is guaranteed to contain no more than one vertex. Since our objective is to resolve between objects of different labels, we can proactively prune the initial vertex octree (see figure \ref{fig:pruning}) such that a leaf node can contain multiple vertices as long as they all have the same label. \red{Clarify this passage: We remove internal binary radix tree nodes seperating identically labeled vertices. For each binary radix tree leaf in parallel, we first look at the label of the leaf's corresponding vertices. If the colors match, we apply that color to the leaf. Otherwise, we mark that leaf as "required".} This initial coloring can be done immediately after the Karras binary radix tree construction without the need to invoke an additional kernel. Before generating the binary radix tree's corresponding quadtree, we propagate the binary radix leaf colors up the tree in parallel, marking internal nodes as "required" when the colors of two children nodes mismatch. Finally, we generate quadtree nodes from only the required internal binary radix tree nodes. By pruning the initial octree, we simplify conflict detection and reduce our memory footprint.

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:pruning-1}
    \includegraphics[width=0.35\columnwidth]{pruned-vs-unpruned/4.PNG} }
  \subfloat[][]{
    \label{fig:pruning-2}
    \includegraphics[width=0.35\columnwidth]{pruned-vs-unpruned/5.PNG} }
  \caption{
    \protect\subref{fig:pruning-1} The initial quadtree built on the object vertices, in which no quadtree cell contains more than one vertex, can be far more complex than needed to resolve between objects.
    \protect\subref{fig:pruning-2} After pruning the quadtree. Quadtree cells can contain multiple vertices as long as they all have the same label.
  }
  \label{fig:pruning}
\end{figure}

@@

Let the ``quadtree address'' refer to the unique ID of a quadtree cell $C$ found by concatenating the local addresses of its ancestors from Root to $C$, where the local address is a 2-bit (3-bit in 3D) Morton code. The address of the root cell is defined as the empty string. Figure \ref{fig:conflict-find-4} shows the address of each leaf cell in a quadtree.

@@

Next we sort the \texttt{BCells} and \texttt{FacetMap} arrays on the \texttt{BCell} values using a parallel radix lexicographical sort (figure \ref{fig:data-structures-2}).% \texttt{BCells} array construction and sorting is done in parallel with the initial Karras octree construction.

\begin{figure}
  \centering
  \begin{tikzpicture}
    \node[anchor=south west,inner sep=0,outer sep=0] at (0,0) {
%      \subfloat[][]{
%        \label{fig:conflict-find-1}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-1.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-2}
        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-4}
        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:conflict-find-5}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
%      \subfloat[][]{
%        \label{fig:conflict-find-6}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
    };

    \node[anchor=south west,inner sep=0,outer sep=0] at (0,-4) {
%      \subfloat[][]{
%        \label{fig:conflict-find-2}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-2.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:conflict-find-4}
%        \includegraphics[width=0.44\columnwidth]{conflict-find-4.pdf} } \hfill
      \subfloat[][]{
        \label{fig:conflict-find-5}
        \includegraphics[width=0.44\columnwidth]{conflict-find-5.pdf} }
      \subfloat[][]{
        \label{fig:conflict-find-6}
        \includegraphics[width=0.44\columnwidth]{conflict-find-6.pdf} }
    };

    %% single column
    %% small rectangle
    %\draw[red,thick] (4.0,1.7) rectangle (6.55,4.2);
    %% upper-right line
    %\draw[red,thin] (6.55,4.2) -- (11.8,4.2);
    %% lower-left line
    %\draw[red,thin] (4.0,1.7) -- (8.0,0.3);
    %% large rectangle
    %\draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);

%    % spanning column, 0.44 size
%    % small rectangle
%    \draw[red,thick] (1.0,1.7) rectangle (6.55,4.2);
%    % upper-right line
%    \draw[red,thin] (6.55,4.2) -- (11.8,4.2);
%    % lower-left line
%    \draw[red,thin] (4.0,1.7) -- (8.0,0.3);
%    % large rectangle
%    \draw[red,thick] (8.0,0.3) rectangle (11.8,4.2);

%    % two column, .22 size
%    % small rectangle
%    \draw[red,thick] (2.0,1.05) rectangle (3.27,2.25);
%    % upper-right line
%    \draw[red,thin] (3.27,2.25) -- (3.9,2.25);
%    % lower-left line
%    \draw[red,thin] (2.0,1.05) -- (3.9,0.4);
%    % large rectangle
%    \draw[red,thick] (3.9,0.4) rectangle (5.8,2.25);

    % two column, .4 size
    % small rectangle
    \draw[red,thick] (0.0,1.65) rectangle (2.33,4.0);
%    % upper-right line
%    \draw[red,thin] (3.27,2.25) -- (3.9,2.25);
%    % lower-left line
%    \draw[red,thin] (2.0,1.05) -- (3.9,0.4);
%    % large rectangle
%    \draw[red,thick] (3.9,0.4) rectangle (5.8,2.25);
  \end{tikzpicture}
%      \subfloat[][]{
%        \label{fig:data-structures-1}
%        \includegraphics[width=0.38\columnwidth]{data-structures-1.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:data-structures-2}
%        \includegraphics[width=0.38\columnwidth]{data-structures-3.pdf} } \hfill
%      \subfloat[][]{
%        \label{fig:data-structures-3}
%        \includegraphics[width=0.38\columnwidth]{data-structures-4.pdf} }
  \caption{
%    \protect\subref{fig:conflict-find-1} We have three objects, blue, red, and green with facets labeled A-I.
    We have three objects, blue, red, and green with facets labeled A-I.
    \protect\subref{fig:conflict-find-2} Initial vertex quadtree.
    \protect\subref{fig:conflict-find-4} Zoomed-in to the region outlined by red in \protect\subref{fig:conflict-find-2} showing the boundary cell (BCell) computation for each facet. %These pairs are given in figure \protect\subref{fig:data-structures-1}.
    \protect\subref{fig:conflict-find-5} Conflict cells, which intersect more than one object, are highlighted.
    \protect\subref{fig:conflict-find-6} The new quadtree after conflict resolution.
%    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
%    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
%\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:conflict-find}
\end{figure}

\begin{figure}
  \centering
      \subfloat[][]{
        \label{fig:data-structures-1}
        \includegraphics[width=0.32\columnwidth]{data-structures-1.pdf} }
      \subfloat[][]{
        \label{fig:data-structures-2}
        \includegraphics[width=0.32\columnwidth]{data-structures-3.pdf} }
      \subfloat[][]{
        \label{fig:data-structures-3}
        \includegraphics[width=0.32\columnwidth]{data-structures-4.pdf} }
  \caption{
    \protect\subref{fig:data-structures-1} The bounding cells (BCells) are stored in an array initially sorted on facet index (letters are used here for clarity). The quadtree array elements are structures which store child and parent pointers (``C/P'' in the figure).
    \protect\subref{fig:data-structures-2} We sort the BCells array using a parallel radix sort on BCell address for fast indexed access. We then, in parallel on each element of the BCells array, store the BCells/FacetMap indices of the first and last facets in a given octree cell in \texttt{FFacet} and \texttt{LFacet}, respectively.
\protect\subref{fig:data-structures-3} For a given octree cell, we can find all contained facets for use in algorithm \ref{alg:find-conflict-cells}.
  }
  \label{fig:data-structures}
\end{figure}

@@

\subsection{Resolve conflict cells}
\label{sec:resolve}

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:conflict-resolution-x}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-x.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-even}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-even.pdf} }\\
  \subfloat[][]{
    \label{fig:conflict-resolution-all}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-all.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-octree}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-octree.pdf} }\\
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-even}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-even.pdf} }
  \subfloat[][]{
    \label{fig:conflict-resolution-adjacent-all}
    \includegraphics[width=0.35\columnwidth]{conflict-resolution-adjacent-all.pdf} }
  \caption{
    \protect\subref{fig:conflict-resolution-x} A conflict cell with two lines from different objects.
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least one sample (red dots).
    \protect\subref{fig:conflict-resolution-even} Fitting boxes such that any box intersecting both lines contains at least two samples. This ensures that a quadtree built from the samples using Karras' algorithm (panel \protect\subref{fig:conflict-resolution-octree}) will have no leaf cells that intersect both lines, ensuring that the new quadtree is locally free of conflict cells.
  }
  \label{fig:conflict-resolution}
\end{figure}
